SciPy 1.0---Fundamental Algorithms for Scientific Computing in Python

### 2019-03-06 scipy-1.0/paper.tex

% Packages
\usepackage[super]{nth}
\usepackage{rotating}
\usepackage{makecell}
\usepackage{pifont}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{float}
\usepackage{pgfplotstable}
\usepackage[multiple]{footmisc}
\usepackage{booktabs,siunitx}

@@

Python, Cython, Fortran, C and C++ are the programming languages used to
implement scientific algorithms in the SciPy library. An analysis of our code
base using the \texttt{linguist} library\cite{linguistref} provides a 
detailed breakdown as \% composition by programming language in 
SciPy (Table~\ref{tab:linguist}).

\begin{table}
\centering
\begin{tabular}{p{3cm}S[table-format = <0.1]}
\toprule
Language & {Percent}\\
\midrule
Python  & 49.5 \\
Fortran & 25.6 \\
C       & 19.5 \\
Cython  & 3.0  \\
C++     & 2.3  \\[\defaultaddspace]
{\TeX, Matlab, Shell, and Makefile} & <0.5 \\
\bottomrule
\end{tabular}
\caption{Language composition of SciPy codebase: lines of code in each
programming language as determined by the \texttt{linguist} package.
The last row denotes tools used in supporting roles in tests,
building, and documentation.}
\label{tab:linguist}
\end{table}

@@

The SciPy test suite is orchestrated by a continuous integration matrix that
includes posix and Windows (32/64-bit) platforms managed by Travis CI and
appveyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
include code linting with pyflakes and pycodestyle. There are more than $13000$
unit tests in the test suite, which is written for usage with the pytest
runner, and with 87 \% and 45 \% line coverages for Python and compiled
code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}).
Some of the historical components of the code base would still benefit from increased
test coverage. Documentation for the code is automatically built and hosted by
the CircleCI service to facilitate evaluation of documentation changes /
integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
version of the Python language.

\begin{figure}[H]
\centering
\begin{tikzpicture}[]
\pgfplotstableread[column/ver/.style=string type]{
ver totpyline covpyline uncovpyline totcompline covcompline uncovcompline
v0.12.1 33844 25721 8123 273634  142290 131344
v0.13.3 36638 28944 7694 338926  172852 166074
v0.14.1 37301 30587 6714 277643  147151 130492
v0.15.1 38339 30288 8051 304898  158547 146351
v0.16.1 40094 32075 8019 322398  167647 154751
v0.17.1 42566 34478 8088 340903  170452 170451
v0.18.1 44711 36216 8495 353417  173174 180243
v0.19.1 43823 36373 7450 436900  200974 235926
v1.0.0 106878 92984 13894 462574 208158 254416
}\covtable
    \begin{axis}[
    width=\textwidth,
    height=7cm,
    enlargelimits=false,
    ymin=0,
    ymax=6e5,
    stack plots=y,%   
    area style,
    xtick=data,
    ytick={2e5,4e5,6e5},
    xticklabels from table={\covtable}{ver},
    scaled y ticks=false,
    y tick label style={/pgf/number format/.cd, fixed, 1000 sep={}},
    legend entries={Compiled (covered),Compiled (uncovered),Python (covered),Python (uncovered)},
    legend style={draw=none,fill=none, cells={align=left},/tikz/every even column/.append style={column sep=3mm}},
    legend columns=-1,
    legend image code/.code={\fill[#1] (0cm,-0.1cm) rectangle (0.2cm,0.1cm);},
    ]
  \addplot+[fill=blue!50] table[x expr=\coordindex, y=covcompline] {\covtable} \closedcycle;
  \addplot+[fill=blue!10] table[x expr=\coordindex, y=uncovcompline] {\covtable} \closedcycle;
  \addplot+[fill=red!70] table[x expr=\coordindex, y=covpyline] {\covtable} \closedcycle;
  \addplot+[fill=red!30] table[x expr=\coordindex, y=uncovpyline] {\covtable} \closedcycle;
\end{axis}
\end{tikzpicture}
\caption{
% This paragraph of caption won't compile
%Python (red) and compiled (blue) code volume in SciPy over time. 
%Deep-shaded area represents lines of code covered by units tests; 
%light-shaded area represents lines not covered. With the exception
%of the removal of $\approx$ 61,000 lines of compiled code for SciPy 
%v0.14 due to (insert reason), the volume of both compiled and Python
%code has increased between releases, as has the number of lines
%covered by unit tests. SciPy v1.0.0 was composed of 462,574 lines of
%compiled code with test coverage near 45\% and 106,878 lines of Python 
%code with test coverage near 87\%.
The reconstitution of historical
build and test environments was performed using a docker-based approach
\cite{scipy-cov}. Analysis of lines of source code covered is performed 
automatically in our continuous integration suite using the \texttt{pytest-cov}
and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
inclusion of additional vendored code.}
\label{fig:coverage}
\end{figure}

@@

SciPy adopted an official Governance Document on August 3,
2017\cite{SciPyProjectGovernance}.  We have a Benevolent Dictator for Life
(BDFL), Pauli Virtanen, leading the project, following a detailed discussion of
the governance model with the community on the mailing list.  In practice, the
overruling authority of the BDFL is anticipated to be used only when the
Steering Council cannot agree on a matter.  The BDFL is expected to consider
any strong indications that they should step down; although the BDFL may
appoint a successor it is generally expected that the Steering Council will be
consulted on the matter.

The (currently) 18 member Steering Council is chaired by Ralf Gommers and
performs a number of tasks including nominating and voting on new Council
Members, and actively contributing to the progress of the project (could be
code review but also outreach activities, etc.). Generally, one year of sustained
and substantial contributions (not restricted to source code contributions)
that improve the project are the prerequisites for nomination of new Council
Members. Council Members do not get special treatment for the vast majority of
the day to day project activities (code submission / review), though it is
anticipated that their experience with the project will serve as helpful
guidance in many matters.  Council Members have commit rights to the project,
but will typically only incorporate changes when there are no substantive
community objections. The Chair of the Steering Council is responsible for
starting more formal technical reviews of the direction of the project,
ensuring that the composition of the Council remains current, and communicating
/ summarizing any Council activities performed privately to the broader
community. We also describe in some detail how institutional partners may
contribute to the project, but no financial weight from grants or other sources
may be leveraged to circumvent or overrule the technical direction of the
project decided upon by the Steering Committee. The Chair does not have a fixed
term, but is nominated by the Steering Committee and is expected to yield to
substantive claims that stepping down is the appropriate action.

SciPy adopted an official Code of Conduct on October 24, 2017
\cite{SciPyCodeOfConduct}, drawing inspiration from the Apache Foundation Code
of Conduct\cite{ApacheCodeOfConduct}, the Contributor Covenant
\cite{ContributorConvenant}, the Jupyter Code of Conduct \cite{Jupyter_COC},
and the Open Source Guides - Code of Conduct\cite{OSG_COC}. In short, we stated
five specific guidelines: \emph{be open} (to anyone participating in our
community), \emph{be empathetic and patient} (in resolving conflicts), \emph{be
collaborative} (we depend on each other to build up the tools in the library),
\emph{be inquisitive} (identifying issues early on can be helpful to everyone),
and \emph{be careful with wording} (do not harass or exclude).  We outlined a
broad diversity inclusion statement, and provide instructions for contacting
the three members of our Code of Conduct Committee or an external
representative on the NumFOCUS Board of Directors.

>>>>

% Packages
\usepackage[super]{nth}
\usepackage{rotating}
\usepackage{makecell}
\usepackage{pifont}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{float}
\usepackage{pgfplotstable}
\usepackage[multiple]{footmisc}

@@

Python, Cython, Fortran, C and C++ are the programming languages used to
implement scientific algorithms in the SciPy library. An analysis of our code
base using the \texttt{linguist} library\cite{linguistref} provides a 
detailed breakdown as \% composition by programming language in 
SciPy (Figure~\ref{fig:linguist}).

\begin{figure}[H]
    \centering
		\includegraphics[width=\textwidth]{static/composition}
    \caption{The breakdown of programming languages used in the
             SciPy library determined using the linguist library.
    	 Small ($<0.5 \%$) amounts of TeX, Matlab, Shell,
    	 and Makefile are excluded for clarity and mostly
    	 provide supporting roles in tests, building, and
    	 documentation.}
    \label{fig:linguist}
\end{figure}

@@

The SciPy test suite is orchestrated by a continuous integration matrix that
includes POSIX and Windows (32/64-bit) platforms managed by Travis CI and
AppVeyor, respectively. Our tests cover Python versions 2.7, 3.4, 3.5, 3.6, and
include code linting with \texttt{pyflakes} and \texttt{pycodestyle}. There are more than $13,000$
unit tests in the test suite, which is written for usage with the \texttt{pytest}
framework, and with 87 \% and 45 \% line coverages for Python and compiled
code, respectively at the SciPy 1.0 release point (Figure~\ref{fig:coverage}). Documentation for the code is automatically built and hosted by
the CircleCI service to facilitate evaluation of documentation changes /
integrity.  Our full test suite also passes with PyPy3, a just-in-time compiled
version of the Python language.

\begin{figure}[H]
\centering
\begin{tikzpicture}[]
\pgfplotstableread[column/ver/.style=string type]{
ver totpyline covpyline uncovpyline totcompline covcompline uncovcompline
v0.12.1 33844 25721 8123 273634  142290 131344
v0.13.3 36638 28944 7694 338926  172852 166074
v0.14.1 37301 30587 6714 277643  147151 130492
v0.15.1 38339 30288 8051 304898  158547 146351
v0.16.1 40094 32075 8019 322398  167647 154751
v0.17.1 42566 34478 8088 340903  170452 170451
v0.18.1 44711 36216 8495 353417  173174 180243
v0.19.1 43823 36373 7450 436900  200974 235926
v1.0.0 106878 92984 13894 462574 208158 254416
}\covtable
    \begin{axis}[
    width=\textwidth,
    height=7cm,
    enlargelimits=false,
    ymin=0,
    ymax=6e5,
    stack plots=y,%   
    area style,
    xtick=data,
    ytick={2e5,4e5,6e5},
    xticklabels from table={\covtable}{ver},
    scaled y ticks=false,
    y tick label style={/pgf/number format/.cd, fixed, 1000 sep={}},
    legend entries={Compiled (covered),Compiled (uncovered),Python (covered),Python (uncovered)},
    legend style={draw=none,fill=none, cells={align=left},/tikz/every even column/.append style={column sep=3mm}},
    legend columns=-1,
    legend image code/.code={\fill[#1] (0cm,-0.1cm) rectangle (0.2cm,0.1cm);},
    ]
  \addplot+[fill=blue!50] table[x expr=\coordindex, y=covcompline] {\covtable} \closedcycle;
  \addplot+[fill=blue!10] table[x expr=\coordindex, y=uncovcompline] {\covtable} \closedcycle;
  \addplot+[fill=red!70] table[x expr=\coordindex, y=covpyline] {\covtable} \closedcycle;
  \addplot+[fill=red!30] table[x expr=\coordindex, y=uncovpyline] {\covtable} \closedcycle;
\end{axis}
\end{tikzpicture}
\caption{
Python (red) and compiled (blue) code volume in SciPy over time.
Deep-shaded area represents lines of code covered by units tests;
light-shaded area represents lines not covered. With the exception
of the removal of $\approx 61,000$ lines of compiled code for SciPy
v0.14, the volume of both compiled and Python
code has increased between releases, as has the number of lines
covered by unit tests. SciPy v1.0.0 was composed of $462,574$ lines of
compiled code with test coverage near $45 \%$ and $106,878$ lines of 
Python code with test coverage near $87 \%$.\\
The reconstitution of historical
build and test environments was performed using a docker-based approach
\cite{scipy-cov}. Analysis of lines of source code covered is performed 
automatically in our continuous integration suite using the \texttt{pytest-cov}
and \texttt{gcov} libraries. Some decreases in compiled code coverage may reflect
inclusion of additional vendored code.}
\label{fig:coverage}
\end{figure}

@@

SciPy adopted an official Governance Document on August 3, 2017\cite{SciPyProjectGovernance}. A Steering Council, currently composed of 18 members, oversees daily development of the project by contributing code and reviewing contributions from the community. Council Members have commit rights to the project repository, but they are expected to merge changes only when there are no substantive community objections. The Chair of the Steering Council, Ralf Gommers, is responsible for initiating biannual technical reviews of project direction and summarizing any private Council activities to the broader community. The project's Benevolent Dictator for Life (BDFL), Pauli Virtanen, has overruling authority on any matter, but the BDFL is expected to act in good faith and only exercise this authority when the Steering Council cannot reach agreement.

SciPy's official Code of Conduct was approved on October 24, 2017. In summary, there are five Specific Guidelines:
\emph{be open} to everyone participating in our community;
\emph{be empathetic and patient} in resolving conflicts;
\emph{be collaborative}, as we depend on each other to build the library;
\emph{be inquisitive}, as early identification of issues can prevent serious consequences; and
\emph{be careful with wording}.
The Code of Conduct specifies how breaches can be reported to a Code of Conduct Committee, and outlines procedures for the Committee's response. Our Diversity Statement ``welcomes and encourages participation by everyone''.
